---
title: "Coronavirus Visualization"
author: "Sahar Saber"
date: "3/30/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Today, I found a dataset about coronavirus updated daily and because there is so much interest in this topic, I thought I would create some data visualizations using ggplot2!

ggplot2 is a library in R used to create interactive and dynamic data visualizations.

*Note: All images were created by myself using ggplot2.*

## Load Required Data

As mentioned above, I used a dataset about coronavirus found at [ourworldindata](https://ourworldindata.org/coronavirus-source-data) 

```{r include = FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(RColorBrewer)
full_data <- read_csv("full_data.csv")
march_29 <- full_data %>% filter(date == "2020-03-29") %>% arrange(desc(total_cases)) %>% select(location,total_cases, total_deaths)

```
```{r}
head(march_29)
```
Also, a dataframe contains the information required to create Choropleth Maps. 
```{r include=FALSE}
corona_worldwide <-march_29[1,]
march_29_countries <- march_29[-1, ]

```
```{r}
world <- map_data("world")
head(world)
```

## Data Manipulation
- A left join was made

```{r}
world_corona <- world %>% left_join(march_29_countries, by = c("region" = "location"))  %>% arrange(desc(total_cases))
```
```{r include=FALSE}
world_regions_corona <-world_corona %>% select(region, total_cases, total_deaths) %>% distinct()
NA_sum <- sum(is.na(world_regions_corona$total_cases))
NA_regions <- world_regions_corona%>% filter(is.na(world_regions_corona$total_cases))


USA_total_cases <- march_29_countries%>% filter(location=="United States") %>% select(total_cases)
USA_total_deaths <- march_29_countries%>% filter(location=="United States") %>% select(total_deaths)

NA_regions$total_cases <- ifelse(NA_regions$region == "USA", USA_total_cases, NA_regions$total_cases)
NA_regions$total_deaths <- ifelse(NA_regions$region == "USA", USA_total_deaths, NA_regions$total_deaths)

UK_total_cases <- march_29_countries%>% filter(location=="United Kingdom") %>% select(total_cases)
UK_total_deaths <- march_29_countries%>% filter(location=="United Kingdom") %>% select(total_deaths)

NA_regions$total_cases <- ifelse(NA_regions$region == "UK", UK_total_cases, NA_regions$total_cases)
NA_regions$total_deaths <- ifelse(NA_regions$region == "UK", UK_total_deaths, NA_regions$total_deaths)

NA_regions$total_cases <- ifelse(is.na(NA_regions$total_cases), 0, NA_regions$total_cases)
NA_regions$total_deaths <- ifelse(is.na(NA_regions$total_deaths), 0, NA_regions$total_deaths)

world_corona_tidied <- world_corona[complete.cases(world_corona$total_cases,world_corona$total_cases ), ] 
NA_regions_full_data <- world %>% inner_join(NA_regions, by = "region")
world_corona_tidied<- world_corona_tidied %>% rbind(NA_regions_full_data)

world_corona_tidied$total_cases <- as.numeric(world_corona_tidied$total_cases)
world_corona_tidied$total_deaths <- as.numeric(world_corona_tidied$total_deaths)
```
- Changing the continuous variables *total_cases* and *total_deaths* into discrete variables
```{r}
world_corona_tidied <- world_corona_tidied %>% mutate(total_cases_discrete = cut(world_corona_tidied$total_cases, 
                                                                               breaks=c(-Inf,0,500, 1000, 10000,100000, Inf), 
                                                                               labels=c("0","<500", "<1000","<10,000","<100,000", ">100,000")))

world_corona_tidied <- world_corona_tidied %>% mutate(total_deaths_discrete = cut(world_corona_tidied$total_deaths, 
                                                                                 breaks=c(-Inf,0,500, 1000, 2000,3000, Inf), 
                                                                                 labels=c("0","<500", "<1000","<2000","<3000", ">3000")))

```


### Choropleth Maps
And Finally, Choropleth Maps shows the spread of the Coronavirus Globally.
```{r}
ggplot(world_corona_tidied, aes(x = long, y = lat, group = group, fill =total_cases_discrete )) +
  geom_polygon(col = "#000000") +
  scale_fill_brewer(palette = "Reds")+
  theme_void()
```


```{r}

ggplot(world_corona_tidied, aes(x = long, y = lat, group = group, fill =total_deaths_discrete )) +
  geom_polygon(col = "#000000") +
  scale_fill_brewer(palette = "Reds")+
  theme_void() 
```







